A Simplified Spectral Deferred Correction Method for Coupling Hydrodynamics with Reaction Networks and Nuclear Statistical Equilibrium

### 2020-10-23 simplified_sdc_intro/paper.tex
remove composition for NSE test
some paper updates

\newcommand{\Shydro}{{{\bf H}}}
\newcommand{\Hb}{{\bf H}}
\newcommand{\Rb}{{\bf R}}
\newcommand{\Rq}{{\bf R}}
\newcommand{\Adv}[1]{{\left [\boldsymbol{\mathcal{A}} \left(#1\right)\right]}}
\newcommand{\Advt}[1]{{\left [\boldsymbol{\mathcal{\tilde{A}}} \left(#1\right)\right]}}
\newcommand{\Advs}[1]{\boldsymbol{\mathcal{A}} \left(#1\right)}

@@

In \cite{castro_sdc}, we developed second- and fourth-order accurate
methods in space and time for coupling hydrodynamics and nuclear
reaction networks based on spectral deferred corrections (SDC)
methods, and demonstrated the method of a variety of test problems.
The approach presented here is considerably simpler, reusing the main
CTU hydrodynamics construction and a largely similar ODE integration
scheme, making this method easier to add to existing simulation codes.
However, it is restricted to second-order accuracy overall.  We term
this algorithm the ``simplified SDC method''.  We use the
\castro\ hydrodynamics code \cite{castro} for all of our numerical
experiments, and all of the code to reproduce the results in this
paper are in the \castro\ github
repository\footnote{\url{https://github.com/amrex-astro/Castro/}}.

@@

Sometimes it is preferable to work with the primitive variables,
\begin{equation}
\qb = \left ( \begin{array}{c}
  \rho \\
  X_k \\
  \alpha_l \\
  u \\
  p \\
  (\rho e) \\
\end{array} \right )
\end{equation}
Here, the system appears
as:
\begin{equation}
\qb_t + \Ab^\xv(\qb) \qb_x  = \Sq
\end{equation}
with the matrix $\Ab^\xv$ giving the coefficients of the spatial derivatives
of the primitive variables:
\begin{equation}
\Ab^\xv(\qb) = \left ( \begin{array}{cccccc}
    u & 0 & 0 & \rho & 0 & 0 \\
    0 & u & 0 & 0    & 0 & 0 \\
    0 & 0 & u & 0    & 0 & 0 \\
    0 & 0 & 0 & u    & 1/\rho & 0 \\
    0 & 0 & 0 & \Gamma_1 p & u & 0 \\
    0 & 0 & 0 & \rho h & 0 & u
  \end{array} \right )
\end{equation}
where $\Gamma_1$ is an adiabatic index, $\Gamma_1 = d\log p/d\log\rho
|_s$.  Note, the primitive state has two thermodynamic quantities, $p$
and $(\rho e)$, to more efficiently handle the general equation of
state in the Riemann solver, as described in \citet{castro}, but
alternate formulations are possible \citep{colellaglaz:1985}.
The source term vector, $\Sq$, can again be decomposed into hydrodynamic
sources (now in terms of the primitive variables) and reaction terms,
\begin{equation}
  \Sq = \Sqhydro + \Rb(q)
\end{equation}
with
\begin{equation}
\label{eq:prim_sources}
\Rb(q) = \left ( \begin{array}{c}
     0 \\
     \omegadot(X_k) \\
     \omegadot(\alpha_l) \\
     0 \\
     \Gamma_1 p \sigma \Sdot \\
     \rho \Sdot
   \end{array} \right )
\end{equation}
where
\begin{equation}
\sigma \equiv \frac{\partial p/\partial T |_\rho}{\rho c_p \partial p/\partial \rho |_T}
\end{equation}
and $c_p$ is the specific heat at constant pressure, $c_p = \partial
h/\partial T |_p$.  A derivation of this source for the pressure
equation can be found in \cite{ABNZ:III}.  We note that this source is
algebraically identical to that shown in Eq.~25 of \cite{castro}.

@@

    Note that our advection terms are piecewise constant \MarginPar{need to think about this more}
    approximations, so when we integrate our system, their
    contributions will be linear in time.  Since $(\rho X_k)$ sees a
    linear advective term, $\rho$ does as well.  $(\rho \Ub)$ and
    $(\rho E)$ will also have linear contributions in time from
    advection.  We can see that the internal energy will also have a
    linear contribution by recognizing that the kinetic energy term,
    $K = (\rho |\Ub|^2)/\rho$ is linear in time.
    Since the reactions don't affect $\rho$ and $(\rho \Ub)$, we can
    algebraically update these using this piecewise linear behavior:
    \begin{align}
      \rho(t) &= \rho^n + \Adv{\rho}^{n+1/2,(k)} \, t \\
      (\rho \Ub)(t) &= (\rho \Ub)^n + \Adv{\rho \Ub}^{n+1/2,(k)} \, t
    \end{align}

@@

        This starts out the same as the above method:
        \begin{equation}
          \nse{\rho, T, (Y_e)_{\rm in}} \rightarrow \bar{A}^\star, (X_k)^\star, (B/A)^\star, dY_e/dt
        \end{equation}
        but we indicate with a $\star$ that most of those values are
        provisional, and we will seek a better value in the corrector step.
        We only update $Y_e$ from this call:
        \begin{equation}
          (Y_e)_{\rm out} = (Y_e)_{\rm in} + \Delta t \frac{dY_e}{dt}
        \end{equation}
        and then we use the table again, but with the updated $Y_e$:
        \begin{equation}
          \nse{\rho, T, (Y_e)_\out} \rightarrow \bar{A}_\out, (X_k)_\out, (B/A)_\out, dY_e/dt
        \end{equation}
        This corrects the composition so that it is consistent with the updated
        $Y_e$.  We use these values $\bar{A}_\out$, $(X_k)_\out$, and $(B/A)_\out$ to then complete the update, computing the energy as in Eq.~\ref{eq:nse_energy}.

@@

With SDC evolution, when we are in NSE, we need to do the advective
and reactive updates together.  With the tabulated NSE reaction
values, we get instantaneous values of the reaction sources, but the only
time derivative we have is for $Y_e$.  We will
use a basic predictor-corrector update:
\begin{equation}
\Uc^\star = \Uc^n + \Delta t \Adv{\Uc}^{n+1/2} + \Delta t \Rb (\Uc^n)
\end{equation}
Calling the NSE table on the starting state gives use:
\begin{equation}
\nse{\rho^n, T^n, (Y_e)^n} \rightarrow \bar{A}^n, (X_k)^n, (B/A)^n, (dY_e/dt)^n
\end{equation}
To get the reactive source, we need to use the iteratively-lagged information
from $\Ic_q^{n+1/2,(k-1)}$:
\begin{equation}
\Rb(\Uc^n) \approx \left (\begin{array}{c}
  0 \\
  \rho^n \Ics_q(\omegadot(X_k)) \\
  \rho^n \Ics_q(\omegadot(\alpha_l)) \\
  0 \\
  \Ics_q(\rho e) 
\end{array}
\right )
\end{equation}
Since we have $(dY_e/dt)^n$ from the table, we replace the source for $Y_e$ in the above
$\Rb(\Uc^n)$ as:
\begin{equation}
R(\rho Y_e) = \rho (dY_e/dt)^n
\end{equation}
which gives an update for the electron fraction as:
\begin{equation}
(\rho Y_e)^\star = (\rho Y_e)^n + \Delta t \Adv{\rho Y_e}^{n+1/2} + \Delta t \rho (dY_e/dt)^n
\end{equation}
From this prediction, we get $\rho^\star$, $\alpha_l^\star$, and
$e^\star$, which allows us to call the EOS and get an updated
temperature, $T^\star$.
We can then again call the NSE table:
\begin{equation}
\nse{\rho^\star, T^\star, (Y_e)^\star} \rightarrow \bar{A}^{n+1,(k)}, (X_k)^{n+1,(k)}, (B/A)^{n+1,(k)}, (dY_e/dt)
\end{equation}
where we use the outputs to define the new values of the composition variables.
Since the density and momenta don't have reactive sources, we just use the $\Uc^\star$ state
as their new values:
\begin{align}
\rho^{n+1,(k)} &= \rho^\star \\
(\rho u)^{n+1,(k)} &= (\rho u)^\star
\end{align}
We then compute the energy release as
\begin{equation}
       (\rho \Sdot)^{n+1/2} = \left [ \rho^{n+1,(k)} \left ( \frac{B}{A} \right )^{n+1,(k)} -
                            \rho^n \left ( \frac{B}{A} \right )^n \right ] N_A \frac{1}{\Delta t}
\end{equation}\MarginPar{is this right?}
and then do the final update of the total energy:
\begin{equation}
(\rho E)^{n+1,(k)} = (\rho E)^n + \Delta t \Adv{\rho E}^{n+1/2} + (\rho \Sdot)^{n+1/2}
\end{equation}

@@

\begin{deluxetable}{lllllll}
\tablecaption{\label{table:nse_strang_methodI} Convergence ($L_1$ norm) for the NSE convergence
problem using Strang splitting + Method I.}
\tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
           \colhead{rate} & \colhead{$\epsilon_{128\rightarrow 256}$} & 
           \colhead{rate} & \colhead{$\epsilon_{256\rightarrow 512}$}}
\startdata
 $\rho$                      & $4.301 \times 10^{19}$  & 1.348  & $1.690 \times 10^{19}$  & 1.166  & $7.532 \times 10^{18}$  \\
 $\rho u$                    & $1.866 \times 10^{28}$  & 1.429  & $6.929 \times 10^{27}$  & 1.101  & $3.230 \times 10^{27}$  \\
 $\rho v$                    & $1.866 \times 10^{28}$  & 1.429  & $6.929 \times 10^{27}$  & 1.101  & $3.230 \times 10^{27}$  \\
 $\rho E$                    & $1.781 \times 10^{38}$  & 1.053  & $8.585 \times 10^{37}$  & 1.021  & $4.231 \times 10^{37}$  \\
 $\rho e$                    & $1.781 \times 10^{38}$  & 1.053  & $8.583 \times 10^{37}$  & 1.021  & $4.230 \times 10^{37}$  \\
 $T$                         & $6.853 \times 10^{21}$  & 1.044  & $3.325 \times 10^{21}$  & 1.018  & $1.642 \times 10^{21}$  \\
 $\rho X(\isotm{He}{4})$     & $1.070 \times 10^{18}$  & 1.186  & $4.701 \times 10^{17}$  & 1.051  & $2.269 \times 10^{17}$  \\
 $\rho X(\isotm{C}{12})$     & $2.166 \times 10^{12}$  & 1.533  & $7.482 \times 10^{11}$  & 1.189  & $3.283 \times 10^{11}$  \\
 $\rho X(\isotm{O}{16})$     & $1.244 \times 10^{13}$  & 1.396  & $4.727 \times 10^{12}$  & 1.117  & $2.179 \times 10^{12}$  \\
 $\rho Y_e$                  & $2.442 \times 10^{19}$  & 1.442  & $8.990 \times 10^{18}$  & 1.212  & $3.880 \times 10^{18}$  \\
 $\rho \bar{A}$              & $1.055 \times 10^{22}$  & 1.032  & $5.158 \times 10^{21}$  & 1.022  & $2.541 \times 10^{21}$  \\
 $\rho (B/A)$                & $4.512 \times 10^{20}$  & 1.386  & $1.727 \times 10^{20}$  & 1.207  & $7.479 \times 10^{19}$  \\
\enddata
\end{deluxetable}

\begin{deluxetable}{lllllll}
\tablecaption{\label{table:nse_strang_methodII} Convergence ($L_1$ norm) for the NSE convergence
problem using Strang splitting + Method II.}
\tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
           \colhead{rate} & \colhead{$\epsilon_{128\rightarrow 256}$} & 
           \colhead{rate} & \colhead{$\epsilon_{256\rightarrow 512}$}}
\startdata
 $\rho$                      & $4.522 \times 10^{19}$  & 1.659  & $1.432 \times 10^{19}$  & 1.542  & $4.917 \times 10^{18}$  \\
 $\rho u$                    & $1.883 \times 10^{28}$  & 1.681  & $5.873 \times 10^{27}$  & 1.524  & $2.042 \times 10^{27}$  \\
 $\rho v$                    & $1.883 \times 10^{28}$  & 1.681  & $5.873 \times 10^{27}$  & 1.524  & $2.042 \times 10^{27}$  \\
 $\rho E$                    & $5.933 \times 10^{37}$  & 1.613  & $1.940 \times 10^{37}$  & 1.411  & $7.293 \times 10^{36}$  \\
 $\rho e$                    & $5.931 \times 10^{37}$  & 1.613  & $1.939 \times 10^{37}$  & 1.411  & $7.291 \times 10^{36}$  \\
 $T$                         & $9.885 \times 10^{20}$  & 1.207  & $4.281 \times 10^{20}$  & 1.058  & $2.057 \times 10^{20}$  \\
 $\rho X(\isotm{He}{4})$     & $5.086 \times 10^{17}$  & 1.430  & $1.887 \times 10^{17}$  & 1.187  & $8.288 \times 10^{16}$  \\
 $\rho X(\isotm{C}{12})$     & $1.585 \times 10^{12}$  & 1.829  & $4.462 \times 10^{11}$  & 1.433  & $1.652 \times 10^{11}$  \\
 $\rho X(\isotm{O}{16})$     & $8.060 \times 10^{12}$  & 1.706  & $2.470 \times 10^{12}$  & 1.347  & $9.710 \times 10^{11}$  \\
 $\rho Y_e$                  & $2.379 \times 10^{19}$  & 1.678  & $7.431 \times 10^{18}$  & 1.578  & $2.489 \times 10^{18}$  \\
 $\rho \bar{A}$              & $5.141 \times 10^{21}$  & 1.230  & $2.191 \times 10^{21}$  & 1.087  & $1.032 \times 10^{21}$  \\
 $\rho (B/A)$                & $4.149 \times 10^{20}$  & 1.693  & $1.283 \times 10^{20}$  & 1.562  & $4.345 \times 10^{19}$  \\
\enddata
\end{deluxetable}

@@

We presented a simple spectral deferred corrections scheme for coupling
hydrodynamics and reactions.

@@

Even though it has no reactive sources, we include $\rho$ in our conservative state
for the purposes of computing the Jacobian (we denote the conserved state used for the
Jacobian $\Uc^\prime$.
Writing this out for two species, $X_\alpha$ and $X_\beta$, we have
\begin{equation}
\Uc^\prime = \left ( \begin{array}{c} \rho \\ \rho X_\alpha \\ \rho X_\beta \\ \rho E \\ \rho e \end{array} \right )
\end{equation}
We take the intermediate state to be ${\bf w} = (\rho, X_\alpha, X_\beta,
K, T)^\intercal$, where $K$ is the kinetic energy:
\begin{equation}
K = \frac{1}{2} |\Ub|^2
\end{equation}

The Jacobian transformation $\partial \Uc^\prime/\partial {\bf w}$ is:
\begin{equation}
\frac{\partial \Uc}{\partial {\bf w}} = \left (
   \begin{array}{ccccc}
       1 & 0 & 0 & 0 & 0 \\
       X_\alpha & \rho & 0 & 0 & 0 \\
       X_\beta & 0 & \rho & 0 & 0  \\
       \rho e_\rho  + e + K &
                 \rho  e_{X_\alpha} & \rho e_{X_\beta} & \rho &
                 \rho e_T \\
       \rho e_\rho  + e  &
                 \rho  e_{X_\alpha} & \rho e_{X_\beta} & 0 &
                 \rho e_T
     \end{array}\right)
\end{equation}
where we use the following notation for compactness:
\begin{equation}
e_\rho = \dedrd \qquad
e_T = \dedTd \qquad
e_{X_k} = \dedXd
\end{equation}
and write the kinetic energy as $K = |\Ub|^2/2$.  We get the inverse
(computed via SymPy, see the included Jupyter notebook) is:

>>>>

\newcommand{\Shydro}{{{\bf H}}}
\newcommand{\Hb}{{\bf H}}
\newcommand{\Rb}{{\bf R}}
\newcommand{\Rq}{{\bf R}}
\newcommand{\Adv}[1]{{\left [\boldsymbol{\mathcal{A}} \left(#1\right)\right]}}
\newcommand{\Advt}[1]{{\left [\boldsymbol{\mathcal{\tilde{A}}} \left(#1\right)\right]}}
\newcommand{\Advss}[1]{{\left [{\mathcal{{A}}} \left(#1\right)\right]}}
\newcommand{\Advs}[1]{\boldsymbol{\mathcal{A}} \left(#1\right)}

@@

In \cite{castro_sdc}, we developed second- and fourth-order accurate
methods in space and time for coupling hydrodynamics and nuclear
reaction networks based on spectral deferred corrections (SDC)
methods, and demonstrated the method of a variety of test problems.
The approach presented here is considerably simpler, reusing the main
CTU hydrodynamics construction and a largely similar ODE integration
scheme, making this method easier to add to existing simulation codes.
It also extends to adaptive mesh refinement with subcycling in a
straightforward manner, avoiding the complications described in
\citep{mccorquodalecolella} needed to fill ghost cells when using
method-of-lines integration.  However, it is restricted to
second-order accuracy overall.  We term this algorithm the
``simplified SDC method''.  We use the \castro\ hydrodynamics code
\citep{castro} for all of our numerical experiments, and all of the
code to reproduce the results in this paper are in the \castro\ github
repository\footnote{\url{https://github.com/amrex-astro/Castro/}}.

@@

Sometimes it is preferable to work with the primitive variables,
\begin{equation}
\qb = \left ( \begin{array}{c}
  \rho \\
  X_k \\
  \alpha_l \\
  u \\
  p \\
  (\rho e) \\
\end{array} \right )
\end{equation}
Here, the system appears
as:
\begin{equation}
\qb_t + \Ab^\xv(\qb) \qb_x  = \Sq
\end{equation}
with the matrix $\Ab^\xv$ giving the coefficients of the spatial derivatives
of the primitive variables:
\begin{equation}
\Ab^\xv(\qb) = \left ( \begin{array}{cccccc}
    u & 0 & 0 & \rho & 0 & 0 \\
    0 & u & 0 & 0    & 0 & 0 \\
    0 & 0 & u & 0    & 0 & 0 \\
    0 & 0 & 0 & u    & 1/\rho & 0 \\
    0 & 0 & 0 & \Gamma_1 p & u & 0 \\
    0 & 0 & 0 & \rho h & 0 & u
  \end{array} \right )
\end{equation}
where $\Gamma_1$ is an adiabatic index, $\Gamma_1 = d\log p/d\log\rho
|_s$.  Note, the primitive state has two thermodynamic quantities, $p$
and $(\rho e)$, to more efficiently handle the general equation of
state in the Riemann solver, as described in \citet{castro}, but
alternate formulations are possible \citep{colellaglaz:1985}.
The source term vector, $\Sq$, can again be decomposed into hydrodynamic
sources (now in terms of the primitive variables) and reaction terms,
\begin{equation}
  \Sq = \Sqhydro + \Rb(\qb)
\end{equation}
with
\begin{equation}
\label{eq:prim_sources}
\Rb(q) = \left ( \begin{array}{c}
     0 \\
     \omegadot(X_k) \\
     \omegadot(\alpha_l) \\
     0 \\
     \Gamma_1 p \sigma \Sdot \\
     \rho \Sdot
   \end{array} \right )
\end{equation}
where
\begin{equation}
\sigma \equiv \frac{\partial p/\partial T |_\rho}{\rho c_p \partial p/\partial \rho |_T}
\end{equation}
and $c_p$ is the specific heat at constant pressure, $c_p = \partial
h/\partial T |_p$.  A derivation of this source for the pressure
equation can be found in \cite{ABNZ:III}.  We note that this source is
algebraically identical to that shown in Eq.~25 of \cite{castro}.

@@

    Note that our advection terms are piecewise
    constant \MarginPar{need to think about this more} approximations,
    so when we integrate our system, their contributions will be
    linear in time.  Since $(\rho X_k)$ sees a linear advective term,
    $\rho$ does as well.  $(\rho u)$ will also have linear
    contributions in time from advection.  Since the reactions don't
    affect $\rho$ and $(\rho u)$, we can algebraically update these
    using this piecewise linear behavior:
    \begin{align}
      \rho(t) &= \rho^n + \Advss{\rho}^{n+1/2,(k)} \, t \\
      (\rho u)(t) &= (\rho u)^n + \Advss{\rho u}^{n+1/2,(k)} \, t
    \end{align}

@@

        This starts out the same as the above method:
        \begin{equation}
          \nse{\rho, T, (Y_e)_{\rm in}} \rightarrow \bar{A}^\star, (X_k)^\star, (B/A)^\star, dY_e/dt
        \end{equation}
        but we indicate with a $\star$ that most of those values are
        provisional, and we will seek a better value in the corrector step.
        We only update $Y_e$ from this call:
        \begin{equation}
          (Y_e)_{\rm out} = (Y_e)_{\rm in} + \Delta t \frac{dY_e}{dt}
        \end{equation}
        and then we use the table again, but with the updated $Y_e$:
        \begin{equation}
          \nse{\rho, T, (Y_e)_\out} \rightarrow \bar{A}_\out, (X_k)_\out, (B/A)_\out, dY_e/dt
        \end{equation}
        This corrects the composition so that it is consistent with the updated
        $Y_e$.  We use these values $\bar{A}_\out$, $(X_k)_\out$, and $(B/A)_\out$ to then complete the update, computing the energy as in Eq.~\ref{eq:nse_energy}.  \MarginPar{we could imagine looping over the EOS to get an updated T and then the table to get the updated $\bar{A}$, to see if that converges?}

@@

With SDC evolution, when we are in NSE, we need to do the advective
and reactive updates together.  With the tabulated NSE reaction
values, we get instantaneous values of the reaction sources, but the only
time derivative we have is for $Y_e$.  We will
use a basic predictor-corrector update:
\begin{equation}
\Uc^\star = \Uc^n + \Delta t \Adv{\Uc}^{n+1/2} + \Delta t \Rb (\Uc^n)
\end{equation}
Calling the NSE table on the starting state gives use:
\begin{equation}
\nse{\rho^n, T^n, (Y_e)^n} \rightarrow \bar{A}^n, (X_k)^n, (B/A)^n, (dY_e/dt)^n
\end{equation}
To get the reactive source, we need to use the iteratively-lagged information
from $\Ic_q^{n+1/2,(k-1)}$:
\begin{equation}
\Rb(\Uc^n) \approx \left (\begin{array}{c}
  0 \\
  \rho^n \Ics_q(\omegadot(X_k)) \\
  \rho^n \Ics_q(\omegadot(\alpha_l)) \\
  0 \\
  \Ics_q(\rho e) 
\end{array}
\right )
\end{equation}
Since we have $(dY_e/dt)^n$ from the table, we replace the source for $Y_e$ in the above
$\Rb(\Uc^n)$ as:
\begin{equation}
R(\rho Y_e) = \rho (dY_e/dt)^n
\end{equation}
which gives an update for the electron fraction as:
\begin{equation}
(\rho Y_e)^\star = (\rho Y_e)^n + \Delta t \Advss{\rho Y_e}^{n+1/2} + \Delta t \rho (dY_e/dt)^n
\end{equation}
From this prediction, we get $\rho^\star$, $\alpha_l^\star$, and
$e^\star$, which allows us to call the EOS and get an updated
temperature, $T^\star$.
We can then again call the NSE table:
\begin{equation}
\nse{\rho^\star, T^\star, (Y_e)^\star} \rightarrow \bar{A}^{n+1,(k)}, (X_k)^{n+1,(k)}, (B/A)^{n+1,(k)}, (dY_e/dt)
\end{equation}
where we use the outputs to define the new values of the composition variables.
Since the density and momenta don't have reactive sources, we just use the $\Uc^\star$ state
as their new values:
\begin{align}
\rho^{n+1,(k)} &= \rho^\star \\
(\rho u)^{n+1,(k)} &= (\rho u)^\star
\end{align}
We then compute the energy release as
\begin{equation}
       (\rho \Sdot)^{n+1/2} = \left [ \rho^{n+1,(k)} \left ( \frac{B}{A} \right )^{n+1,(k)} -
                            \rho^n \left ( \frac{B}{A} \right )^n \right ] N_A \frac{1}{\Delta t}
\end{equation}\MarginPar{is this right?}
and then do the final update of the total energy:
\begin{equation}
(\rho E)^{n+1,(k)} = (\rho E)^n + \Delta t \Advss{\rho E}^{n+1/2} + (\rho \Sdot)^{n+1/2}
\end{equation}

@@

\begin{deluxetable}{lllllll}
\tablecaption{\label{table:nse_strang_methodI} Convergence ($L_1$ norm) for the NSE convergence
problem using Strang splitting + Method I.}
\tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
           \colhead{rate} & \colhead{$\epsilon_{128\rightarrow 256}$} & 
           \colhead{rate} & \colhead{$\epsilon_{256\rightarrow 512}$}}
\startdata
 $\rho$                      & $4.301 \times 10^{19}$  & 1.348  & $1.690 \times 10^{19}$  & 1.166  & $7.532 \times 10^{18}$  \\
 $\rho u$                    & $1.866 \times 10^{28}$  & 1.429  & $6.929 \times 10^{27}$  & 1.101  & $3.230 \times 10^{27}$  \\
 $\rho v$                    & $1.866 \times 10^{28}$  & 1.429  & $6.929 \times 10^{27}$  & 1.101  & $3.230 \times 10^{27}$  \\
 $\rho E$                    & $1.781 \times 10^{38}$  & 1.053  & $8.585 \times 10^{37}$  & 1.021  & $4.231 \times 10^{37}$  \\
 $\rho e$                    & $1.781 \times 10^{38}$  & 1.053  & $8.583 \times 10^{37}$  & 1.021  & $4.230 \times 10^{37}$  \\
 $T$                         & $6.853 \times 10^{21}$  & 1.044  & $3.325 \times 10^{21}$  & 1.018  & $1.642 \times 10^{21}$  \\
 $\rho Y_e$                  & $2.442 \times 10^{19}$  & 1.442  & $8.990 \times 10^{18}$  & 1.212  & $3.880 \times 10^{18}$  \\
 $\rho \bar{A}$              & $1.055 \times 10^{22}$  & 1.032  & $5.158 \times 10^{21}$  & 1.022  & $2.541 \times 10^{21}$  \\
 $\rho (B/A)$                & $4.512 \times 10^{20}$  & 1.386  & $1.727 \times 10^{20}$  & 1.207  & $7.479 \times 10^{19}$  \\
\enddata
\end{deluxetable}

\begin{deluxetable}{lllllll}
\tablecaption{\label{table:nse_strang_methodII} Convergence ($L_1$ norm) for the NSE convergence
problem using Strang splitting + Method II.}
\tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
           \colhead{rate} & \colhead{$\epsilon_{128\rightarrow 256}$} & 
           \colhead{rate} & \colhead{$\epsilon_{256\rightarrow 512}$}}
\startdata
 $\rho$                      & $4.522 \times 10^{19}$  & 1.659  & $1.432 \times 10^{19}$  & 1.542  & $4.917 \times 10^{18}$  \\
 $\rho u$                    & $1.883 \times 10^{28}$  & 1.681  & $5.873 \times 10^{27}$  & 1.524  & $2.042 \times 10^{27}$  \\
 $\rho v$                    & $1.883 \times 10^{28}$  & 1.681  & $5.873 \times 10^{27}$  & 1.524  & $2.042 \times 10^{27}$  \\
 $\rho E$                    & $5.933 \times 10^{37}$  & 1.613  & $1.940 \times 10^{37}$  & 1.411  & $7.293 \times 10^{36}$  \\
 $\rho e$                    & $5.931 \times 10^{37}$  & 1.613  & $1.939 \times 10^{37}$  & 1.411  & $7.291 \times 10^{36}$  \\
 $T$                         & $9.885 \times 10^{20}$  & 1.207  & $4.281 \times 10^{20}$  & 1.058  & $2.057 \times 10^{20}$  \\
 $\rho Y_e$                  & $2.379 \times 10^{19}$  & 1.678  & $7.431 \times 10^{18}$  & 1.578  & $2.489 \times 10^{18}$  \\
 $\rho \bar{A}$              & $5.141 \times 10^{21}$  & 1.230  & $2.191 \times 10^{21}$  & 1.087  & $1.032 \times 10^{21}$  \\
 $\rho (B/A)$                & $4.149 \times 10^{20}$  & 1.693  & $1.283 \times 10^{20}$  & 1.562  & $4.345 \times 10^{19}$  \\
\enddata
\end{deluxetable}

@@

We presented a simplified spectral deferred corrections scheme for coupling
hydrodynamics and reactions.

Future work is to extend this methodology to MHD.

@@

Even though it has no reactive sources, we include $\rho$ in our conservative state
for the purposes of computing the Jacobian (we denote the conserved state used for the
Jacobian $\Uc^\prime$.
Writing this out for two species, $X_\alpha$ and $X_\beta$, we have
\begin{equation}
\Uc^\prime = \left ( \begin{array}{c} \rho \\ \rho X_\alpha \\ \rho X_\beta \\ \rho E \\ \rho e \end{array} \right )
\end{equation}
We take the intermediate state to be ${\bf w} = (\rho, X_\alpha, X_\beta,
K, T)^\intercal$, where $K$ is the kinetic energy:
\begin{equation}
K = \frac{1}{2} u^2
\end{equation}

The Jacobian transformation $\partial \Uc^\prime/\partial {\bf w}$ is:
\begin{equation}
\frac{\partial \Uc}{\partial {\bf w}} = \left (
   \begin{array}{ccccc}
       1 & 0 & 0 & 0 & 0 \\
       X_\alpha & \rho & 0 & 0 & 0 \\
       X_\beta & 0 & \rho & 0 & 0  \\
       \rho e_\rho  + e + K &
                 \rho  e_{X_\alpha} & \rho e_{X_\beta} & \rho &
                 \rho e_T \\
       \rho e_\rho  + e  &
                 \rho  e_{X_\alpha} & \rho e_{X_\beta} & 0 &
                 \rho e_T
     \end{array}\right)
\end{equation}
where we use the following notation for compactness:
\begin{equation}
e_\rho = \dedrd \qquad
e_T = \dedTd \qquad
e_{X_k} = \dedXd
\end{equation}
and write the kinetic energy as $K = u^2/2$.  We get the inverse
(computed via SymPy, see the included Jupyter notebook) is:
